---
title: "Coursera - Regression Models Assignment"
author: "Arkadiusz Oliwa"
date: "20 march 2018"
output: 
      html_document: 
            keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(corrplot)
library(formattable)

```

## Executive summary

In this assignment I compared manual transmission and automatic transmission for MPG (Miles/US galon). There was used simple exploratory analysis as well as hypothesis testing and linear regression. As you can see in this paper having a manual car cause increase in MPG in comparison to automatic transmission.

## Dataset

The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles

```{r, warning=FALSE, message=FALSE}
mtcars %>%
  formattable(align="l") %>%
  as.datatable()
```


```{r, echo=FALSE}
mtcars$am <- as.factor(mtcars$am)
```


## Simple exploratory data analysis

Table  and box plot show car transmission types by  MPG. As you can see there is increase in MGP for manual transmission in comparison to automatic transmission

```{r, warning=FALSE, message=FALSE}
mtcars %>%
  mutate(Transmission = ifelse(am == 0, no = "manual", yes = "automatic")) %>%
  group_by(Transmission) %>%
  summarise(median = median(mpg), mean = mean(mpg), sd = sd(mpg), min = min(mpg), max = max(mpg)) %>%
  formattable(align="l")
```

```{r}
mtcars %>%
  mutate(am = ifelse(am == 0, no = "manual", yes = "automatic")) %>%
  ggplot(aes(y = mpg, x = am, fill = am)) +
  geom_boxplot(alpha = .7, varwidth = TRUE) 
```

We can also see  that there is linear correlation between mpg and the variables disp, hp, wt (strong negative) and drat (strong positive)  

```{r}
corrplot(corr = cor(mtcars[, c("mpg", "disp", "hp", "drat", "wt", "qsec")], 
                    method = "pearson"), 
         method = "number", type = "lower")
```


## Hypothesis testing

I’m interested in if the average value of MPG  differs significantly from a manual and  automatic transmission within a defined confidence level 0.05

```{r}
t.test(formula = mpg ~ am, data = mtcars)
```

The p-value is 0.001374 so we reject the null hypothesis and we inference that there is a significant statistical difference in the mean MPG between manual transmission cars and that of automatic transmission cars

## Linear regression

I’m going to use function step() to automatically choose the best model by AIC criterion in a stepwise algorithm


```{r, echo=TRUE, warning=FALSE, message=FALSE, results='hide'}
fit <- step(object = lm(formula = mpg ~ ., data = mtcars), direction = "both")
```



Based on diagnostic plots we can say that residual are normally distributed and homoskedastic. 

```{r}
par(mfrow=c(2,2))
plot(fit)
```


```{r}
summary(fit)
```

Looking at result we can say that:
The adjusted R-squared for the model is 0.8336, and the three variables coefficients are all significant at the 5% confidence level. 
Based on coefficients we can say that having a manual car we gain 2.9358 MPG above that of an automatic 






